What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-07 android-stacks.tex
A bit more on RQ2

\input{android-stack-stats}

@@

\noindent \fbox{
\begin{minipage}{0.96\columnwidth} 
\emph{RQ2: Most exceptions are due to programming mistakes
or generally irrecoverable errors.\todo{Revise this}}
\end{minipage}}

@@

To better understand the main causes of such exceptions we evaluated semantics
related to the top 100 root causes reported on stack-issues of both Java dataset
and Android subset (which correspond to approximately 77\% of all exceptions
reported reported on Java stack-issues and 95\% of all exceptions reported on
Android stack-issues), and classified them according to the the categories
presented in Table~\ref{tophundrend}.

@@

As we can see from Table 4, the that programming errors are the causes of most
of the exception stack traces reported on issues. In the Android subset we could
find more issues related to Backward compatibility and security issues than on
the general data set.

@@

Concerning the size of the stacks (i.e. the number of frames composing them), we
could observe that  sometimes it exceeded 100 frames (158 stack frames).
Manually inspecting the corresponding execution trace was caused by one of the
two reasons: recursive calls; exceptions signaled by a method defined deep down
in a reused framework; exceptions successively wrapped. Regardless the reason,
in exception flows involving many methods, it is piratically impossible to
handle the exception a way different than existing the  application, because
there is hardly enough contextual information to perform any other recovery
action.

%TODO: Include a Histogram here.

@@

@@

%\noindent\emph{3) Odd Wrappings}
\subsection{Odd Wrappings}

Table~\ref{wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions where indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace (represented as the table lines
on which the root cause is none). This analysis also revealed unexpected
wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and
Error; and an Error wrapping a Runtime.

@@

In this section, we present works that are related to our own, distributed in
four categories: i) works that use the information available on stack traces;
ii) empirical studies on the usage of Java exceptions and its fault proneness;
iii) tools to extract stack traces information from natural language artifacts
(i.e., issues and emails), and iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
employed dataflow analysis to find the propagation paths of checked and
unchecked exception types. Modeled after Robillard and Murphy's work, other
tools have been proposed to support the static analysis of exception
flows~\cite{coelho2008assessing}. The main limitation off all
static analysis tools is the number of false positives inherent to static analysis
solutions, which can lead to a high number of exception flows, specially if
considering Java Environment exceptions and exceptions signaled from libraries.

Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the combined use of stack traces extracted from issues and bytecode and source
code analysis.

@@

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
investigated the degree of reuse across applications in Android Market, the
study showed that almost 23\% of the classes inherited from a base class in the
Android API, and that 217 mobile apps were reused completely by another mobile
app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
discussions of Android platform and found out that approximately 20\% of
energy-related bugs in Android occurred after an OS update. McDonnell et
al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
Android API and 10 dependent applications using the version history data found
in github. The study found that approximately 25\% of all methods in the client
code used the Android API, and that the methods reusing fast-evolving APIs were
more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
approximately 7K free Android apps and observed that the last successful apps
used Android APIs that were on average 300\% more change-prone than the APIs
used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
reports of 5 Android applications an observed that 29\% had to do with poor
exceptional handling code. Our work differs from the others as it aims at
distilling the information of bug reports describing uncaught exceptions created
for mobile applications in Github, in order to get a first view
of what is causing the crashes across applications available and what the
characteristics of the stack traces can tell us about the exception structure of
those applications.

>>>>

@@

@@

To better understand the main causes of such exceptions, we evaluated semantics
related to the top 100 root causes reported on stack-issues of both Java dataset
and Android subset (which correspond to approximately 77\% of all exceptions
reported reported on Java stack-issues and 95\% of all exceptions reported on
Android stack-issues), and classified them according to the the categories
presented in Table~\ref{tab:tophundrend}.

@@

Programming errors are the causes of most of the exception stack traces reported
on issues. In the Android subset we could find more issues related to Backward
compatibility and security issues than on the general data set.

@@

Concerning the size of the stacks (i.e. the number of frames composing them), we
could observe that  sometimes it exceeded 100 frames (158 stack frames).
Manually inspecting the corresponding execution trace was caused by one of the
two reasons: recursive calls or exceptions signaled by a method defined deep
down in a reused framework; exceptions successively wrapped. Irrespective of the
reason, in exception flows involving many methods, it is practically impossible
to handle an exception in a way other than exiting the application, as at the
place the exception occurs there is hardly enough contextual information to
perform recovery actions.

@@

\noindent \fbox{
\begin{minipage}{0.96\columnwidth} 
\emph{RQ2: Most exceptions are due to programming mistakes
or generally irrecoverable errors.}
\end{minipage}}

@@

%\noindent\emph{3) Odd Wrappings}
\subsection{Odd Wrappings}
Table~\ref{wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions where indeed wrapped in runtime exceptions or even errors, while most
of them were not wrapped along the stack trace (represented as the table lines
on which the root cause is none). This analysis also revealed unexpected
wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and
Error; and an Error wrapping a Runtime.

@@

In this section, we present works that are related to our own, distributed in
four categories: i) works that use the information available on stack traces;
ii) empirical studies on the usage of Java exceptions and its fault proneness;
and iii) tools to extract stack traces information from natural language artifacts
(i.e., issues and emails)
%, and iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
employed dataflow analysis to find the propagation paths of checked and
unchecked exception types. Modeled after Robillard and Murphy's work, other
tools have been proposed to support the static analysis of exception
flows~\cite{coelho2008assessing}. The main limitation off all
static analysis tools is the number of false positives inherent to static analysis
solutions, which can lead to a high number of exception flows, specially if
considering Java Environment exceptions and exceptions signaled from libraries.
Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
exception handling code of 32 open-source systems, both for Java and .NET. They
observed that the action handlers were very simple (e.g., logging and present a
message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
set of bad practices on exception handling that hinder software maintainability
and robustness, based on their own experience with Java enterprise applications.
Our work differs from those two works as we tried to identify bad practices from
the combined use of stack traces extracted from issues and bytecode and source
code analysis.

@@

%\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
%investigated the degree of reuse across applications in Android Market, the
%study showed that almost 23\% of the classes inherited from a base class in the
%Android API, and that 217 mobile apps were reused completely by another mobile
%app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
%discussions of Android platform and found out that approximately 20\% of
%energy-related bugs in Android occurred after an OS update. McDonnell et
%al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
%Android API and 10 dependent applications using the version history data found
%in github. The study found that approximately 25\% of all methods in the client
%code used the Android API, and that the methods reusing fast-evolving APIs were
%more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
%approximately 7K free Android apps and observed that the last successful apps
%used Android APIs that were on average 300\% more change-prone than the APIs
%used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
%reports of 5 Android applications an observed that 29\% had to do with poor
%exceptional handling code. Our work differs from the others as it aims at
%distilling the information of bug reports describing uncaught exceptions created
%for mobile applications in Github, in order to get a first view
%of what is causing the crashes across applications available and what the
%characteristics of the stack traces can tell us about the exception structure of
%those applications.

### 2014-05-07 stacktrace-stats.tex
A bit more on RQ2

% latex table generated in R 2.15.3 by xtable 1.7-1 package
% Mon Apr 28 15:44:10 2014
\begin{table*}[ht]
\centering
\begin{tabular}{rp{27em}rrrrc}
  \hline
Feature & Description & quant\_5 & mean & median & quant\_95 & histogram \\ 
  \hline
all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
  with\_exception & All issues featuring an exception. & 0.00 & 3.21 & 0.00 & 6.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-0c1305fc40f695dbdd1f4696db24f147.pdf} \\ 
  with\_stack & All issues featuring a stack trace. & 0.00 & 2.67 & 0.00 & 4.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-896524802e915aa3f01ef1ecd2489e7e.pdf} \\ 
  defect\_issues & All issues labeled as defect. & 0.00 & 5.51 & 0.00 & 12.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d0ab95cd3e300c21b0a40d6bec58e4c.pdf} \\ 
  defect\_with\_exception & All issues labeled labeled as defect and including an exception. & 0.00 & 1.82 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-502028d1a7e2e77414e2f9f63e0bcd50.pdf} \\ 
  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 1.73 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-dab0764147debd99066c84814d46d50e.pdf} \\ 
   \hline
\end{tabular}
\caption{Descriptive statistics for the analyzed issue dataset. Historgrams are in log scale.} 
\label{tab:stacktrace-stats}
\end{table*}

>>>>

% latex table generated in R 2.15.3 by xtable 1.7-1 package
% Mon Apr 28 15:44:10 2014
\begin{table*}[ht]
\centering
\begin{tabular}{rp{27em}rrrrc}
  \hline
Feature & Description & quant\_5 & mean & median & quant\_95 & histogram \\ 
  \hline
   \multicolumn{2}{l}{\bfseries{Java}}  \\
all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
  with\_exception & All issues featuring an exception. & 0.00 & 3.21 & 0.00 & 6.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-0c1305fc40f695dbdd1f4696db24f147.pdf} \\ 
  with\_stack & All issues featuring a stack trace. & 0.00 & 2.67 & 0.00 & 4.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-896524802e915aa3f01ef1ecd2489e7e.pdf} \\ 
  defect\_issues & All issues labeled as defect. & 0.00 & 5.51 & 0.00 & 12.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d0ab95cd3e300c21b0a40d6bec58e4c.pdf} \\ 
  defect\_with\_exception & All issues labeled labeled as defect and including an exception. & 0.00 & 1.82 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-502028d1a7e2e77414e2f9f63e0bcd50.pdf} \\ 
  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 1.73 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-dab0764147debd99066c84814d46d50e.pdf} \\ 
   \hline
   \multicolumn{2}{l}{\bfseries{Android}}\\
   all & All issues recorded for the project. & 2.00 & 67.70 & 20.00 & 248.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d4a87daa44c63097df9bb1f3506d541.pdf} \\ 
  with\_exception & All issues featuring an exception. & 1.00 & 8.16 & 2.00 & 34.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-bf73235d0503f937ef6f36c28cf1ba0d.pdf} \\ 
  with\_stack & All issues featuring a stack trace. & 1.00 & 6.44 & 2.00 & 27.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-528cb6be67074e6e2c22d11245920e13.pdf} \\ 
  defect\_issues & All issues labeled as defect. & 0.00 & 13.28 & 1.00 & 52.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-eb1797e86ef23b2a07907e5ccbeff7b9.pdf} \\ 
  defect\_with\_exception & All issues labeled labeled as defect and including an exception. & 0.00 & 3.25 & 0.00 & 13.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-f1a879616aa3e0995db35c6c02bc8520.pdf} \\ 
  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 2.92 & 0.00 & 10.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-49dbfa02c366cab8b489f33014de587d.pdf} \\
 \hline
 \end{tabular}
\caption{Descriptive statistics for the analyzed issue dataset. Historgrams are in log scale.} 
\label{tab:stacktrace-stats}
\end{table*}
